Kit Ligand and Kit receptor tyrosine kinase sustain synaptic inhibition of Purkinje Cells

Version 1

Abstract

Summary

The cell-type specific expression of ligand/receptor and cell-adhesion molecules is a fundamental mechanism through which neurons regulate connectivity. Here we determine a functional relevance of the long-established mutually exclusive expression of the receptor tyrosine kinase Kit and the trans-membrane protein Kit Ligand by discrete populations of neurons in the mammalian brain. Kit is enriched in molecular layer interneurons (MLIs) of the cerebellar cortex (i.e., stellate and basket cells), while cerebellar Kit Ligand is selectively expressed by a target of their inhibition, Purkinje cells (PCs). By in vivo genetic manipulation spanning embryonic development through adulthood, we demonstrate that PC Kit Ligand and MLI Kit are required for, and capable of driving changes in, inhibition of PCs. Collectively, these works in mice demonstrate that the Kit Ligand/Kit receptor dyad sustains mammalian central synapse function and suggest a rationale for the affiliation of Kit mutation with neurodevelopmental disorders.

Introduction

The proto-oncogene receptor tyrosine kinase Kit (c-Kit, CD117) is an evolutionarily conserved, loss-of-function intolerant, gene that is enriched in specific neuron populations and tentatively associated with rare cases of neurological dysfunction (pLI= 0.98 and LOEUF of 0.17) (Figure S1A,B)1,2. It is perhaps unsurprising that loss-of-function Kit mutations in humans are rare; Kit is highly pleiotropic, supporting diverse cell populations including melanocytes, hematopoietic stem/progenitor cells, germ cells, masts cell, and interstitial cells of Cajal3. Thus, Kit mutations can result in diverse conditions such as hypo- or hyperpigmentation or melanoma, anemia or leukemia, infertility, mastocytosis, and impaired gut motility or gastro-intestinal tumors (Kit biology reviewed 2 4).

The activation of Kit is stimulated by Kit Ligand (Gene symbol is KITLG, herein KL), a single-pass transmembrane protein having an extracellular active domain that induces Kit dimerization and kinase activity. Across several decades and species, it has been reported that KL and Kit are mutually expressed in discrete neuron populations known to be connected; it has thus been hypothesized that the expression pattern of KL-Kit may reflect a role in connectivity5-8. Consistent with such a function, case reports have implicated inactivating Kit mutations with neurological impairment (References9-13 and Figure S1C). Clinical phenotypes include developmental delay, ataxia, hypotonia, intellectual disability, deafness, and autism spectrum disorder. White matter abnormalities have been noted, and limited evidence suggests KL-Kit may indeed regulate axon outgrowth. KL or Kit mutation or pharmacological manipulation alters outgrowth of spinal commissural axons in vitro14, KL stimulates cortical neurite outgrowth in vitro 15, and Kit reduction in cortex of developing rats or mice delays axon extension16. While these reports provide anatomical/morphological data suggesting that KL-Kit influences connectivity, critically, these studies did not address functional consequences to neuronal physiology or synaptic connectivity. Functional studies of KL-Kit in synaptic physiology have likely been hampered by pleiotropy and organismal or cell death. Severe global KL-Kit hypomorphs exhibit fatal anemia, and developmental Kit depletion or inhibition in the cerebrum has resulted in death of neuronal progenitors or nascent neurons16-18. Therefore, whether KL-Kit is necessary for synaptic function has remained largely unstudied. Here, we address this gap.

In mice, rats, and humans, Kit is abundantly expressed by molecular layer interneurons (MLIs) of the cerebellar cortex6-8,19, while cerebellar KL is restricted to Purkinje Cells (PCs), which MLIs provide GABAergic synaptic inhibition to. We tested the hypothesis that PC KL and MLI Kit were essential to the GABAergic inhibition of PCs. We created a Kit conditional knockout mouse and accomplished embryonic knockout of Kit from all MLIs; separately, we knocked out KL from postnatal PCs. By either method, disruption of the KL-Kit ligand-receptor pair produced robust and specific impairments to GABAergic inhibition of PCs. Through sparse postnatal viral manipulation of KL in PCs, we provide evidence that KL functions throughout adulthood to regulate synaptic function. These results demonstrate that cell-type specific expression of KL and Kit influences functional connectivity in the mammalian brain, informing the long-observed expression of KL-Kit in the brain and suggesting a rationale for the affiliation of Kit loss with neurological impairment.

Methods

Animals

All procedures performed at Michigan State University were approved by the Institutional Animal Care and Use Committees of Michigan State University, which is accredited by the Association for Assessment and Accreditation of Laboratory Animal Care (AAALAC). Mice were on a 12-hour dark/light cycle, with ad libitum food and water. Male and female mice were used for all experiments. Mice were genotyped in-house using standard PCR-based methods, or via Transnetyx.

Creation of Kit conditional knockout mouse

We generated a strain of mice having exon 4 of Kit floxed. This was accomplished by utilizing the services of UNC Animal Models Core (Dr. Dale Cowley) to generate chimeras from an ES cell line having a Knockout-first Allele (with conditional potential) for Kit (HEPD0509_8). The ES clones were obtained through the EuMMCR (European Mouse Mutant Cell Repository) of the EUCOMM (European Conditional Mouse Mutagenesis Program)20. Chimeric mice were screened for germline transmission of the targeted allele and were then crossed to a Flp deleter strain (C57BL/6N-Albino-Rosa26-Flpo) to convert the knockout-first allele to a conditional allele (tm1c) through removal of the FRT-flanked selection cassette. F1 male offspring carrying Kit tm1c and Flp alleles were crossed to C57Bl6/J females; we selected resultant progeny heterozygous for Kit tm1c but negative for Flp and the selection cassette. The Kit tm1c alle was bred to homozygosity while selecting against animals carrying the Tyr c-Brd allele (previously conveyed by C57BL/6N-Albino-Rosa26-Flpo). The strain of mice having the Kit floxed tm1c allele are given the nomenclature C57BL/6N–Kittm1c(EUCOMM)Mirow/J; the Mirow lab code is provisional. Additional strains of mice used were Pax2-Cre transgenic mice (STOCK Tg(Pax2-cre)1Akg/Mmnc, RRID:MMRRC_010569-UNC21, Kit eGFP (Tg(Kit-EGFP)IF44Gsat/Mmucd; Gensat) 22 backcrossed for greater than 5 generations to C57Bl6/J (Stock No: 000664, The Jackson Laboratory), KitL floxed (Kitltm2.1Sjm/J, Stock #017861) 23, and Pcp2-Cre (Stock # 004146) mice 24, all from The Jackson Laboratory.

To produce mixed litters of “Control” and “Kit KO” animals, mice homozygous for the Kit tm1c allele but negative for Pax2-Cre were crossed to mice homozygous for the floxed allele and Pax2-Cre positive. We used Kit KO females crossed to Control males as our breeding scheme. Experimental animals were derived from at least 5 generations of backcrossing Kit KO animals to Kit tm1c homozygous animals derived from a separate (Kit tm1c homozygous x Kit tm1c homozygous) breeding cohort.

Validation of Kit knockout

We investigated conditional knockout of Kit protein from the cerebellum by Western blot of total protein lysates from acutely harvested cerebellums (∼P52) and by indirect immunofluorescence on free-floating vibratome-generated sections of transcardial-perfused formaldehyde-fixed mouse brains (∼P31). Western blots were probed with Rabbit: α-Kit (clone D13A2, Cell Signaling Technologies product 3074), α-GAPDH (Clone 14C10, product 2118 Cell Signaling Technology), or α-parvalbumin antibody (product PV-27, Swant), each primary was detected via HRP conjugated Goat α-Rabbit secondary antibody (Biorad, product 170-6515) after incubation with Clarity ECL substrate (Bio-Rad, 1705061) via a BioRad ChemiDoc MP.

For indirect immunofluorescence, primary antibodies were rat α-Kit (Rat α-Kit clone ACK4, ThermoFisher, product MA5-17836), mouse α-calbindin (α-calbindin-D28K, CB-955, Sigma Aldrich), and rabbit α-parvalbumin (PV-27, Swant). Primaries were detected by Goat-host highly cross-adsorbed secondary antibodies coupled to Alexa 488, Cy3, or Alexa 647 (Jackson Immunoresearch). Nuclei were labeled with DAPI, tissues were mounted in an anti-fade reagent, and fluorescent signals captured laser scanning confocal microscopy. Cell densities for MLIs (parvalbumin positive and/or calbindin negative somas of the molecular layer) were performed on Maximum Z-projections of 2.5-2.6 micron thick stacks, using the multi-point feature for counting somas, and the polygon ROI tool for determining the area of the Molecular Layer in which the somas were counted; results from two-technical replicates were averaged for each animal. The polygon tool was also used for determining maximal PSD95+ pinceau area. To determine the width of the molecular layer, seven linear ROIs per field of view of comparable locations of lobules IV/V were averaged to generate per animal layer thickness measurements. For analysis of synapses, we quantified puncta in the molecular layer that were triple positive for rabbit α-VGAT(131 003), mouse α-Gephyrin (147 011), and Chicken α-Calbindin (214 006), all from Synaptic Systems. For analysis of pinceau markers we utilized mouse α-Kv1.1 (K36/15), Kv1.2 (K14/16), and PSD-95 (K28/43), all from NeuroMab/AntibodiesInc.

Stereotaxic Procedures

Injections of replication-defective viral particles used protocols substantially similar to those previously published 25. In brief, mice were brought to an anesthetic plane using inhaled isoflurane by a low-flow system (SomnoSuite, Kent); thermal regulation was assisted by heating pad, and pain managed through topical lidocaine and ketoprofen. A digital stereotaxic device (Kopf) and reference atlas were used to localize a small hole in the skull; cerebellar coordinates were: neonates; A/P, -2.55; M/L, - 1.1 (reference Lambda), D/V, 1.5 to 0.5 from dura; juvenile; A/P, -6.25; M/L, - 1.35; D/V, 1.5 to 0.75 (reference Bregma); adults; A/P, -6.35; M/L, - 1.8; D/V, 1.5 to 0.75 (reference Bregma). Viral particles were delivered via a ∼30-gauge Hamilton Syringe controlled by digital syringe pump (WPI). Virus volume was 1-2 ul and there was a 5-minute period after infusion prior to needle withdrawal. Animals were monitored during and after surgical operations. Ages for injection were, nominally, postnatal day (P) 7, 19, and 56; for P7 and P19, variability of up to two days was utilized to accommodate differences in animal size, and for P56 the eighth week of life was utilized.

Viruses

Lentiviral particles were derived from vectors previously described to express transgenes under the hUbiC promoter and pseudotyped with VSV-G 26,27. Silent mutations disrupted the internal EcoRI sites of mouse KitL (MC204279, Origene), the KitL coding sequence was then flanked by EcoRI sites, and the subsequent KitL coding sequence was introduced into a FUC-T2A-(EcoRI) plasmid, to create a FUC-T2A-KL-1 plasmid, wherein C designates mCherry. Subsequent deletion of KitL Exon 6 generated a KL-2 vector 28-31. Sequences were verified by sequencing. Adeno-Associated Viral (AAV) particles were created by Vigene to drive Cre expression by the Purkinje Cell-Specific L7.6 promoter 32, or to drive mCherry or mCherry-T2A-KL-2 in a Cre-On fashion under the EF1α promoter, with the AAV1 serotype.

Electrophysiology

For patch clamp electrophysiology of Control vs Kit KO, cells were recorded in slices generated from animals at 36.9 ± 0.6 and 37.4 ± 0.9 days for MLIs and PCs respectively. Avertin-anesthetized mice were perfused with a carbogen-equilibrated ice-cold slicing solution containing (in mM): 110 C5H14ClNO, 7 MgCl2.6H2O, 2.5 KCl, 1.25 NaH2PO4, 25 NaHCO3, 0.5 CaCl2-2H2O, 10 Glucose, and 1.3 Na-Ascorbate. In the same solution, we generated 250-micron thick parasagittal sections of the cerebellum (Leica VT1200). Slices recovered at 34°C for 30 minutes in a carbogenated ACSF containing (in mM): 125 NaCl, 25 NaHCO3, 1.25 NaH2PO4, 2.5 KCl, 1 MgCl2.6H2O, 1 CaCl2-2H2O and 25 Glucose; slices were then held at room temperature for at least 30 minutes before recording. Recordings were performed in carbogenated external recording ACSF (32.7 ± 0.1°C) containing (in mM): 125 NaCl, 25 NaHCO3, 1.25 NaH2PO4, 2.5 KCl, 1 MgCl2.6H2O, 2 CaCl2-2H2O and 25 Glucose. MLIs and PCs of folia IV/V were targeted for recording by IR-DIC, or epifluorescence stimulated by a CoolLED pE-4000, detected via a SciCam Pro on a SliceScope Pro 6000-based rig (Scientifica). Recording electrodes were pulled (Narishige, PC-100) from standard-wall borosilicate glass capillary tubing (G150F-4, Warner Instruments) and had 5.1 ± 0.08 and 2.8 ± 0.02 MΩ tip resistance for MLIs and PCs respectively. Spontaneous Inhibitory Postsynaptic Currents (sIPSCs) and Miniature Inhibitory Postsynaptic Currents (mIPSC) were recorded with an intracellular solution containing (in mM):140 CsCl, 4 NaCl, 0.5 CaCl2-2H2O, 10 HEPES, 5 EGTA, 2 Mg-ATP, and 0.4 Na-GTP, 2 QX-314. For action potential recordings and for Miniature Excitatory Postsynaptic Currents (mEPSCs), the intracellular solution contained (in mM): 140 K-gluconate, 10 KCl, 1 MgCl2, 10 HEPES, 0.02 EGTA, 3 Mg-ATP, and 0.5 Na-GTP as described previously 33. The internal pipette solution pH was adjusted to 7.35 with CsOH for IPSCs and mIPSCS, and with KOH for mEPSCs and action potentials, while for all the osmolarity was adjusted to 300 mOsmolL-1 with sucrose. In whole-cell voltage clamp mode PCs were held at -70 mV; to isolate sIPSCs, NBQX (10 μM) and D-AP5 (50 μM) were added to the recording solution and mIPSCs recordings additionally included 1 μM tetrodotoxin (TTX). To record mEPSCs, 1 μM TTX and 50 μM picrotoxin was added to extracellular solution. Action potentials were recorded in MLIs within the middle third of the molecular layer; spontaneous action potentials were recorded in I=0 mode, or action potentials were evoked by depolarizing current injection in 600 ms steps of 10 pA, from 0 to 150 pA, with a 7 second inter-sweep interval. For paired pulse, current ranging from 150 uA to 250uA was delivered by concentric bipolar stimulating electrode in the molecular layer. Stimulus duration was 50 microseconds, and ISI was 50 mS. For each cell, 10 sweeps were recorded with 10 S inter sweep interval. Purkinje Cells with an access resistance of 10-20 (15.9 ± 0.5), or MLIs with an access resistance <30 (23.6 ± 0.6) MΩ were considered for recording. Recordings with >20% change in series resistance were excluded from analysis. Signals were acquired at 10 KHz with a low-noise data acquisition system (Digidata 1550B) and a Multiclamp700-A amplifier and were analyzed using pClamp11.1 (Molecular Devices). Both mIPSCs and mEPSCs with over 8 pA amplitude and longer than 1 ms duration were included while overlapping events were rejected from analysis; also rejected from analysis was any miniature event for which amplitude, rise, and decay metrics were not all available. Analysis of the PCP2 Cre mediated KL KO tissues was conducted similarly at ∼40-44 days old. For viral PC KL KO animals injected as neonates (P7), juvenile (P19), and adults (P56) were recorded at an average of ∼35, 45, and 83 days old, respectively.

Software

Microscopic image processing was conducted via FIJI/ImageJ, electrophysiological data was analyzed by pClamp11.1, statistical analyses were conducted via GraphPad Prism 9, and Fig.s were created with BioRender, PowerPoint, and GraphPad Prism 9.

Results

We generated mice in which Kit exon 4 is flanked by LoxP sites (Kit tm1c); Kit gene modifications are illustrated in Figure 1A, B. Kit tm1c homozygous Control animals had normal appearance, sex ratios, and reproductive success. Kit is abundantly expressed in GABAergic interneurons in the molecular layer of the cerebellar cortex (MLIs), whereas Kit Ligand (KL) is expressed by Purkinje cells (PCs). To determine if Kit influences synapse function, we conditionally knocked out Kit from MLIs and assessed the MLI-PC synapse. Pax2 is a transcription factor expressed early in the lineage of GABAergic interneurons of the cerebellum34; a Pax2-Cre transgene enables embryonic recombination35,36. We generated Kit tm1c homozygous litters of which nominally half were hemizygous for Pax2 Cre (Kit KO) and half were not (Control) Figure 1C. We produced Control and Kit KO animals in normal sex and genotype ratios. As expected for the C57Bl6/J background, Control littermates had dark coat and eyes, but littermate Kit KO animals had white whiskers, white fur with variable pigmented patches, and black eyes; example littermates Figure 1D. Given Pax2 expression in melanocytes37 and the role of Kit in melanogenesis (Reviewed in38,39), the pigmentation phenotype provided gross confirmation of conditional Kit KO. We thus confirmed that cerebella of Kit KO animals were Kit depleted. By immunohistochemistry in Control animals, we recapitulated the known pattern of Kit immunoreactivity in the parvalbumin positive MLIs of the cerebellar cortex (1F and Inset). In contrast, a sex-matched Kit KO littermate demonstrated loss of cerebellar Kit immunoreactivity. By Western blot of cerebellar lysates from Control animals, Kit resolved as a major band at ∼120 kD, but in Kit KO littermates, Kit immunoreactivity was lost; GAPDH as loading control was equivalent, as was parvalbumin (which marks mature MLIs and PCs) Figure 1G.

Design and validation of a Kit conditional knockout mouse.

(A,B) Kit receptor tyrosine kinase (Kit) is enriched in parvalbumin positive GABAergic interneurons of the molecular layer (i.e. basket and stellate cells, MLIs) of the cerebellar cortex, where they synapse upon each other and upon Purkinje cells (PCs, Calbindin+), which express Kit Ligand (KL). Scale bar 10 microns. Expression pattern schematized in B.

(C) In humans and mice, Kit is encoded by up to 21 exons, which in mouse is encoded on the plus strand of chromosome 5 at 75,735,647-75,817,382 bp. We generated a Kit conditional knockout mouse in which Kit Exon 4 is flanked by LoxP sites.

(D) We generated Control mice homozygous for the Kit floxed allele Kittm1c(EUCOMM)Mirow which varied in Pax2-Cre)1Akg transgene status, with the goal of depleting Kit from MLIs in embryonic development.

(E) Male and Female, Control and Kit KO mice were born in equal ratios. KO mice were notably hypopigmented in hair and whiskers, though not eyes.

(F) Confocal microscopy of Kit immunoreactivity in age and sex matched Control and Kit KO littermates demonstrates the established enrichment of Kit in the molecular layer of the cerebellar cortex, and its loss in Kit KO. Scale Bar 500 microns top row, 50 microns inset.

(G) Utilizing a distinct assay and primary antibody, we confirm the detection of Kit immunoreactivity in Controls, and its loss in Kit KO litter mates of either sex by Western Blot of total protein lysates of cerebella. We affirmed equivalent protein loading by GAPDH and parvalbumin.

Having validated that Kit KO animals lacked cerebellar Kit, we assessed the functional impact. We compared spontaneous GABAergic inhibitory post synaptic currents (sIPSC) in PCs of acute slices generated from ∼P34 Control and Kit KO animals, schema Figure 2A, example traces Figure 2B. Kit KO was associated with a significant change in the distribution of sIPSC event amplitudes (Figure 2C). Calculating average sIPSC event frequency, amplitude, and duration for each PC of Control or Kit KO revealed a significant decrease in sIPSC frequency (1D) but not amplitude or duration. The reduced sIPSC frequency in Kit KO was not associated with changes in density or distribution of MLIs or in thickness of the molecular layer (Figure S2).

The knockout of Kit from Cerebellar Cortex Interneurons Impairs Inhibition of Purkinje Cells.

(A,B) Experimental design and example traces of postsynaptic currents in Purkinje Cells (PCs) from Control animals or from those with Kit KO from cerebellar cortex molecular layer interneurons (MLIs). Scale bar is 500 ms x 100 pA.

(C) A frequency distribution plot for individual sIPSC event amplitudes recorded in PCs as in A,B. A KS test reveals a significant difference in the distribution of these event amplitudes; p<0.0001, n in chart.

(D) For each PC, the average Frequency, Amplitude, and Duration of sIPSC was determined. There was a ∼50% decrease in sIPSC frequency p=0.0013, but no change in Amplitude or Duration.

(E-G) Experimental design: In separate experiments, we recorded miniature postsynaptic currents from PCs or from MLIs in Control and in Kit KO animals.

(H) Example traces of mIPSCs or (I) mEPSCs recorded in PCs, and example traces of mIPSC in MLIs, all from Control or from Kit KO animals. Scale for H and J is 50 pA, 25 pA for I, all 500 ms.

(K-M) Analysis of per cell average miniature event Frequency and Amplitude revealed that Kit KO reduced mIPSC frequency, but not amplitude, in PCs by >50%, p=0.013. mEPSCs in PCs and mIPSC in MLIs were not different between Control and Kit KO.

N in charts, refers to number of cells. Error bars are SEM. Statistical significance is by two-tailed t-test, with Welch’s correction as needed. NS indicates p≥0.05.

We sought to determine if reduced PC inhibition was related to altered MLI physiology (Schema, Figure S3A). Capacitance, input resistance, and spontaneous or stimulated MLI action potential frequency was not different between Control and Kit KO (Figure S3B-E). Since MLIs were present and firing normally in Kit KO, we sought to determine if there were defects in evoked neurotransmitter release. Direct electrode stimulation of the outer or inner molecular layer in a paired-pulse paradigm (Schema, Figure S3F) evoked IPSCs of equivalent amplitude in PCs from Control and Kit KO, and we detected no difference in the paired pulse ratio (Figure S3G,H). These data suggest that MLI synapses could function normally in Kit KO under direct stimulation. We therefore investigated markers of the MLI:PC synapses to determine if they were altered.

The density of GABAergic synapses upon PCs (triple positive VGAT, Gephyrin, and Calbindin puncta in the molecular layer) was not reduced by Kit KO (example immunofluorescence, Figure S4A). However, GABAergic puncta upon PCs were significantly smaller (∼10%) in Kit KO (Figure S4B). MLI axon terminals contribute not just to individual axo-dendritic synaptic puncta, but also to interdigitated MLI axon collaterals around the soma and initial axon segment of PCs, forming “pinceaux” structures. By parvalbumin immunoreactivity, these structures were preserved but smaller in Kit KO (arrowheads Figure S4C), and this was confirmed by decreased Kv1.2 (Figure S4D) and PSD95 (Figure S4D). As PSD95 immunoreactivity faithfully follows multiple markers of pinceaux size 40, we quantified PSD95 immunoreactive pinceau area and determined that pinceaux area was decreased by ∼50% in Kit KO (n 26 Control vs 43 Kit KO, p<0.0001, two-tailed t-test). These results suggested that Kit KO MLIs may have defects in physical and/or functional maturation of synapses upon PCs.

To evaluate the number of functional GABAergic synapses, we analyzed miniature synaptic currents in acute cerebellar slices, (schema Figure 2 E,F,G; example traces Figure 2 H,I,J). In PCs of Kit KO, there was a significant ∼50% reduction in the average frequency, but not amplitude, of mIPSCs (Figure 2K). In contrast, the average frequency and amplitude of mEPSCs in PCs from Control and Kit KO did not differ (Figure 2L). These data suggested that Kit KO impaired the MLI:PC synapse but did not rule out a general defect in MLI synapse function. As MLIs synapse not just upon PCs, but also upon MLIs, we evaluated mIPSCs in MLIs. Mean mIPSC frequency and amplitude was unchanged between Control and Kit KO (Figure 2M). Therefore, embryonic MLI Kit KO was associated with a specific defect in the number or proportion of functional GABAergic synapses upon PCs. As MLI Kit and PC KL expression is maintained postnatally, we hypothesized that postnatal postsynaptic PC KL KO would phenocopy embryonic presynaptic MLI Kit KO.

We utilized previously described KL floxed23 and PC-Cre24 strains to yield mice with Control or KL KO PCs (Schema Figure 3A). We confirmed depletion of cerebellar KL transcripts by rtPCR in data not shown23, and we confirmed that PC KL KO did not disrupt the pattern of Kit immunoreactivity. Patch clamp of PCs in acute slices revealed that PC KL KO produced significant differences in the distribution of sIPSC event amplitudes (Example traces Figure 3B, Figure 3C). PC KL KO reduced the average sIPSC frequency, amplitude, and duration in PCs (Figure 3D). The phenotype was specific to GABAergic synapses: there was a substantial reduction in the frequency but not the amplitude of mIPSC events in PCs (Figure 3E). In contrast, frequency and amplitude of mEPSCs in KL KO PCs was not different (Figure 3F). In data not shown, a KS test revealed a significant difference in the individual mEPSC event amplitude distribution (p=0.0002, n=1196 Control vs 1244 KL KO), however the mean mEPSC amplitude was within 1% between conditions. Similarly, a KS test of mIPSC event amplitudes revealed a significant difference (p<0.0001, n=5996 Control vs 3958 KL KO), but the mean mIPSC amplitude was ∼10% greater in PC KL KO than in Control. It is therefore unlikely that decreased event amplitude contributed to the specific and robust >50% decrease in mIPSC frequency detected in PC KL KO. Thus, like MLI Kit KO, PC KL KO reduced inhibition of PCs.

Knockout of Kit Ligand from Purkinje Cells decreases their inhibitory input.

(A,B) Experimental design and example traces of inhibitory postsynaptic currents detected in Purkinje Cells (PCs) from Control animals or from those with Kit Ligand knockout (KL KO) accomplished by a PCP2-Cre x KL Floxed strategy. Scale bar is 500 ms x 100 pA.

C) A frequency distribution plot for individual sIPSC event amplitudes recorded in PCs as in A,B. A KS test reveals a significant difference in the distribution of these event amplitudes; p<0.0001, n in chart.

(D) For each PC, the average Frequency, Amplitude, and Duration of sIPSC was determined. There was a ∼30-40%% decrease in sIPSC Frequency p=0.0078, Amplitude p=0.024 and Duration p=0.0006.

(E, F) The average mIPSC recorded in PC KL KO vs Control PCs had a >50% decreased Frequency p<0.0001 without a corresponding decrease in mIPSC Amplitude. The average mEPSC Frequency and Amplitude recorded in separate PCs did not differ between Control and KL KO.

N in charts, refers to number of cells. Error bars are SEM. Statistical significance is by two-tailed t-test, with Welch’s correction as needed. NS indicates p>=0.05.

Recombination by PCP2 Cre begins at ∼P724, whereas Pax2 Cre recombination occurs at ∼E15 21,34. Our results thus suggest that continued expression of the KL-Kit dyad sustains inhibitory drive to PCs. To test this hypothesis (and avoid transformative effects of Kit overexpression), we tuned KL postnatal expression in vivo.

To determine if acute focal manipulations in KL modulate PC inhibition, we injected replication-defective viruses leveraging the PC-specific L7.6 promoter32 to express Cre in KL floxed animals, and we identified Cre-positive PCs by mCherry expression (Schema Figure 4A, Example transduction, Figure S5A-C,F-H). Injections were at P7, P18, or P56; acute slices were generated 2-3 weeks later. Patch-clamp recordings in areas of sparse infection revealed that regardless of when KL was depleted, PC KL KO reduced PC sIPSC frequency (P7, P18, and P56 in Figure S5, P18 injections Figure 4C,D). We next examined PC KL over-expression. We co-injected L7.6 Cre AAV with AAV encoding EF1α driven Cre-on mCherry and the membrane bound isoform of mouse KL (mKL2) via a T2A element to produce PC KL overexpression (OX) (Schema Figure 4B). Whether at P18 or P56, KL OX PCs had increased sIPSC frequency and amplitude compared to nearby Control PCs (Figure 4E,F).

Local levels of Kit Ligand influence the inhibition Purkinje cells receive.

(A) To accomplish in vivo mixtures of Control and Kit Ligand knockout (KL KO) Purkinje Cells (PCs), an AAV encoding Cre under the PC specific L7.6 promoter was co-injected with an AAV encoding a Cre-On mCherry cassette under the Ef1α promoter.

(B) To accomplish in vivo mixed population of Control and Kit Ligand overexpressing (KL OX) PCs, the L7.6 Cre AAV was co-injected with AAV expressing mCherry and (T2A) murine Kit Ligand isoform 2 under the Ef1α promoter.

(C,D) In animals injected at P18, PC KL KO neurons demonstrated a ∼70% decrease in sIPSC event frequency compared to adjacent Control PCs (p<0.0001) or PCs recorded from Sham control animals (p=0.0002). Control vs Sham sIPSC frequency was not significantly different. Similarly, the sIPSC event amplitude was significantly reduced by PC KL KO vs adjacent Control PCs (p=0.0009) or those from Sham animals (p=0.31). Control vs Sham sIPSC amplitude was not different.

(E, F) In animals injected at P18, PC KL OX neurons demonstrated an 8-fold increase in sIPSC Frequency vs local Control PCs (p<0.0001). The frequency of sIPSC events in KL OX PCs was not significantly different from PCs recorded in different Sham animals; this Sham PC sIPSC frequency (while comparable across studies) was 7-fold higher than Control PCs within the KL OX experimental animals (p<0.0001). A similar pattern was found for sIPSC amplitude, being significantly elevated (3-fold) in the KL OX PCs vs their neighboring Control PCs (p=0.0003), but not distinguishable from those recorded in Sham animals.

(G) Proposed model for a function of Purkinje Cell Kit Ligand and Molecular Layer Interneuron Kit. Under normal conditions, the balanced expression of PC KL mediates equivalent presynaptic function of MLI axons via Kit. In focal PC KL KO or PC KL OX, those PCs with the locally highest KL levels sustain nominal levels of MLI mediated inhibition at the expense of reduced functional input to out-competed PCs.

N in charts, refers to number of cells. Error bars are SEM. Statistical significance is by Brown-Forsythe ANOVA test with Dunnett’s T3 multiple comparisons test. NS indicates p>=0.05.

We noted that Control PCs from the PC KL KO paradigm had sIPSC frequency and amplitude similar to PCs from Sham injections (Figure 4C,4D). In contrast, while KL OX PCs received elevated GABAergic input compared to their adjacent Control PCs, this (KL OX) relative increase in inhibitory drive was not beyond the inhibition recorded in PCs from separate Sham animals (Figure 4F). That is, the focal gain in synaptic input to KL OX PCs appears to come at the expense of decreasing input to nearby Control PCs. Thus, KL-Kit levels may not set the number of MLI:PC release sites, but may instead influence the proportion of synapses that are functional for neurotransmission (Figure 4G). Such a requirement is consistent with the impaired synaptic function observed under both widespread conditional knockout and sparse viral manipulation.

Discussion

For genes essential to survival or reproduction, germline knockout strategies are not feasible for evaluating postnatal physiology. Among such genes, Kit is an example of a highly pleiotropic gene, and so the interpretation of global hypomorphs is challenging and the recombination of conditional alleles outside of desired tissues can lead to lethal or sterile phenotypes. By generating a conditional knockout mouse, we were able to produce viable animals with Kit depleted from the cerebellum to reveal a role in synapse function.

MLI number, distribution, action potential firing, and the density of GABAergic synaptic puncta upon PCs, are all normal in MLI Kit KO. These data suggest that while MLI axon terminals are present in Kit KO (perhaps related to the smaller size of synaptic puncta and pinceaux), they may be less functionally mature. An increased failure rate between action potential generation and neurotransmitter release would be consistent with the normal firing rate and VGAT puncta density but decreased sIPSC and mIPSC frequency observed. Under the PPR paradigm, the apparent equivalent evoked responses may be due to direct stimulation of axon terminals and/or recruitment of multiple MLIs.

Kit KO most directly impacts the MLI:PC synapse, based upon normal miniature events for PF/CF:PC (mEPSC in PC) and MLI:MLI (mIPSC in MLI) synapses. This is perhaps expected given the cell type specific expression of KL by PCs and Kit by MLIs, however, it is a critical distinction as Cerebellins41 and Calsystenin42 impact multiple forms of synaptic input to PCs. It will be informative to determine if the residual inhibitory drive to PCs observed in MLI Kit KO or PC KL KO comes from PC axon collaterals43-46 whose input we might expect to be preserved if KL-Kit functions primarily to mediate connectivity between different cell types.

Our results demonstrate that the KL-Kit axis, from embryonic development through young adulthood (P56), can alter GABAergic drive to PCs. These results suggest that KL-Kit may be necessary to maintain, and perhaps capable of modulating, connectivity in the adult cerebellum. As each MLI (e.g. basket cell) may have axon collaterals that target multiple PCs, and as each PC may receive input from multiple MLIs, it is interesting to speculate that dynamics in the expression or shedding of KL may act through presynaptic Kit to tune the degree of convergent MLI:PC inhibitory drive. This would complement existing forms of modulation at the MLI:PC synapse, such those dependent upon other PC derived signaling molecules, like endocannabinoids47-49 or peptides like secretin50,51 and CR 52. The physiology of PCs, and the structure of MLI:PC inhibition, also varies over zones such as those identified by Zebrin-II and PLC β453-57; whether there is anatomical variation in KL or Kit expression, signaling, or dependency, remains to be determined.

Kit is a receptor tyrosine kinase that can activate multiple downstream effectors including Src family kinases, PLC, PI3K/mTOR, and the MAPK pathway, the latter two notable for their shared affiliation with synapse phenotypes and autism spectrum disorder. Furthermore, trans-cellular interactions between KL and Kit may function in a kinase-independent cell adhesion molecule modality58,59. There is a substantial literature illustrating the importance of adhesion molecules in the organization and maintenance of MLI:PC inhibition, such as neurexins/neuroligins, dystrophin/dystroglycan, semaphorins, and neurofascins41,42,60-64. Much of what is understood about the maintenance of PC inhibition is revealed by post-synaptic phenotypes; that is, cis-and trans interacting proteins that, via gephyrin, maintain the organization of PC GABAA receptors. It is notable therefore that MLI Kit or PC KL KO have a more predominant effect on mIPSC frequency than on amplitude, suggesting a presynaptic phenotype. A compatible model is that PC derived KL may act retrograde/trans-synaptic to modulate MLI axon function via Kit binding/activity.

Despite the reduced PC inhibition resulting from MLI Kit KO or from PC KL KO, we have not observed overt cerebellar signs (i.e., ataxia). We suspect that this is due to compensation. Even the developmental ablation of functional GABAA channels in PCs does not produce overt motor signs, while adulthood ablation ca 65. Thus, we might unmask behavioral phenotypes if KL or Kit KO is delayed. Beyond direct inhibition of PCs, MLI inhibition constrains granule cell mediated PC dendritic excitation and thus gates the conversion of PC LTD to LTP at the PF-PC synapse in motor learning66,67. It is therefore possible that KL or Kit KO may only reveal phenotypes in motor learning (or other cerebellar-plasticity dependent) behaviors. Not inconsistent with a role in learning, global hypomorphs for KL or Kit have altered hippocampal neurophysiology and Water Maze performance68-70. Continuous expression of KL and Kit in specific neurons may thus ultimately reflect not only a role in development or maintenance of synaptic function as suggested here, but also in plastic modulation of synaptic strength.

Despite the importance of future works to determine molecular mechanisms and organismal consequences, our study is a discrete advance because it establishes that KL-Kit regulate mammalian central synapse function. Here, we demonstrate impacts at the mouse GABAergic MLI:PC synapse; however, Kit and KL exist in diverse organisms and neuronal populations including glutamatergic neurons of the hippocampus, and neurons in olfactory bulb, basal forebrain, and brainstem. It is therefore feasible that the synaptic phenotypes reported here reflect that KL-Kit is broadly capable of influencing connectivity.

eLife Assessment

This valuable study from Zaman et al. demonstrates that the cKit-Kit ligand complex is necessary for the formation and/or maintenance of molecular layer interneuron synapses in cerebellar Purkinje cells. The evidence presented is solid; in particular, the use of cell-type specific knockout of cKit in molecular layer interneurons and knockout of Kit ligand in Purkinje cells provides robust evidence. This work will be of particular relevance to those interested in inhibitory synapse formation or the role of inhibition in Purkinje cell behavior.

Reviewer #1 (Public Review):

This manuscript from Zaman et al., investigates the role of cKit and Kit ligand in inhibitory synapse function at molecular layer interneuron (MLI) synapses onto cerebellar Purkinje cells (PC). cKit is a receptor tyrosine kinase expressed in multiple tissues, including select populations of neurons in the CNS. cKIt is activated by Kit ligand, a transmembrane protein typically expressed at the membrane of connected cells. A strength of this paper is the use of cell-specific knockouts of cKit and Kit ligand, in MLIs and PCs, respectively. In both cases, the frequency of spontaneous or miniature (in the presence of TTX) IPSCs was reduced. This suggests either a reduction in the number of functional inhibitory release sites or reduced release probability. IPSCs evoked by electrical stimulation in the molecular layer showed no change in paired-pulse ratio, indicating release probability is not changed in the cKit KO, and favoring a reduction in the number of release sites. Changes in IPSC amplitude were more subtle, with some analyses showing a decrease and others not. These data suggest that disruption of the cKit-Kit ligand complex reduces the number of functional synapses with only minor changes in synapse strength. However, immunolabelling of inhibitory synapses in cKit KO mice using VGAT and Gephyrin antibodies revealed no change in the number of puncta, but reduced size of puncta. This result is more consistent with reduced synapse strength (size) without a change in synapse number. The apparent contradiction of these results is not resolved. It would be interesting to know if immunolabeling of inhibitory synapses in Kit ligand KO mice would produce similar results.

In separate experiments, the authors used viral expression of Cre (driven by the PC-specific L7 promotor) to sparsely KO Kit ligand in PCs. In recordings from neighboring Cre+ (Kit ligand KO) and Cre - (kit ligand intact) PCs, the spontaneous IPSC frequency and amplitude were reduced. Using a similar viral approach, they also overexpressed Kit ligand in wild-type PCs. Here the results are more difficult to interpret. The frequency and amplitude of spontaneous IPSCs were significantly greater in Cre+ PCs compared to Cre- cells. However, the effect appears to be primarily due to a drastic reduction in IPSC amplitude and frequency in the control Cre- cells rather than an increase in Cre+ cells. This puzzling result is interpreted as evidence that cKit influences the proportion of synapses that are functional for neurotransmission, but not the number of release sites. Though this interpretation is not described in detail.

Overall, this paper makes great use of genetic and viral approaches to examine the function of cKit and Kit ligand at MLI-PC synapses. Measurements are generally limited to immunolabeling and spontaneous IPSC recordings, a wider variety of approaches, such as EM imaging or recording from connected MLI-PC pairs would likely provide more detail on specific pre- or postsynaptic phenotypes and more clearly determine whether the number or strength of synapses is changing.

Reviewer #2 (Public Review):

In their study, Zaman et al. demonstrate that deletion of either the receptor tyrosine kinase Kit from cerebellar interneurons or the kit ligand (KL) from Purkinje cells reduces the inhibition of Purkinje cells. They delete Kit or KL at different developmental time points, illustrating that Kit-KL interactions are not only required for developmental synapse formation but also for synapse maintenance in adult animals. The study is interesting as it highlights a molecular mechanism for the formation of inhibitory synapses in Purkinje cells.

The tools generated, such as the floxed Kit mouse line and the virus for Kit overexpression, may have broader applications in neuroscience and beyond.

However, to enhance the publication's impact and strengthen its hypotheses, conclusions, and scientific rigor, it would be beneficial to include additional experimental details, data analyses (particularly regarding the quantification of electrophysiology data), as well as methodological and textual clarifications.

One general weakness is that Kit expression is not limited to molecular layer interneurons but also extends to the Purkinje layer and Golgi interneurons. Although this expression may not conflict with the reported results, as Purkinje layer interneurons form few or no synapses onto Purkinje cells, it should be highlighted in the text (introduction and/or discussion).

In summary, the data support the hypothesis that the interaction between Kit and KL between cerebellar Molecular Layer Interneurons and Purkinje Cells plays a crucial role in promoting the formation and maintenance of inhibitory synapses onto PCs. This study provides valuable insights that could inform future investigations on how this mechanism contributes to the dynamic regulation of Purkinje cell inhibition across development and its impact on mouse behavior.

Reviewer #3 (Public Review):

Summary:

Bidirectional transsynaptic signaling via cell adhesion molecules and cell surface receptors contributes to the remarkable specificity of synaptic connectivity in the brain. Zaman et al., investigate how the receptor tyrosine kinase Kit and its trans-cellular kit ligand regulate molecular layer interneuron (MLI)- Purkinje cell (PC) connectivity in the cerebellum. Presynaptic Kit is specific for MLIs, and forms a trans-synaptic complex with Kit ligand in postsynaptic PC cells. The authors begin by generating Kit cKOs via an EUCOMM allele to enable cell-type specific Kit deletion. They cross this Kit cKO to the MLI-specific driver Pax2-Cre and conduct validation via Kit IHC and immunoblotting. Using this system to examine the functional consequences of presynaptic MLI Kit deletion onto postsynaptic PC cells, they record spontaneous and miniature synaptic currents from PC cells and find a selective reduction in IPSC frequency. Deletion of Kit ligand from postsynaptic PC cells also results in reduced IPSC frequency, together supporting that this trans-synaptic complex regulates GABAergic synaptic formation or maturation. The authors then show that sparse Kit ligand overexpression in PCs decreases neighboring uninfected control sIPSCs in a potentially competitive manner.

Strengths:

Overall, the study addresses an important open question, the data largely support the authors' conclusions, the experiments appear well-performed, and the manuscript is well-written. I just have a few suggestions to help shore up the author's interpretations and improve the study.

Weaknesses:

The strong decrease in sIPSC frequency and amplitude in control uninfected cells in Figure 4 is surprising and puzzling. The competition model proposed is one possibility, and I think the authors need to do additional experiments to help support or refute this model. The authors can conduct similar synaptic staining experiments as in Fig S4 but in their sparse infection paradigm, comparing synapses on infected and uninfected cells. Additional electrophysiological parameters in the sparse injection paradigm, such as mIPSCs or evoked IPSCs, would also help support their conclusions.

The authors should validate KL overexpression and increased cell surface levels using their virus to support their overexpression conclusions.

Author Response

The following is the authors’ response to the original reviews.

Reviewer #1 (Recommendations for The Authors):

Major comments:

1. The immunolabeling data in Figure S4 shows no change in puncta number but reduced puncta size in Kit KO. sIPSC data show reduced frequency but little change in amplitude. These data would seem contradictory in that one suggests reduced synaptic strength, but not number, and the other suggests reduced synapse number, but not strength. How do the authors reconcile these results?

Regarding the synaptic puncta, In Kit KO (or KL KO), we have not detected an overt reduction in the average VGAT/Gephyrin/Calbindin positive puncta density or puncta size per animal. With respect to puncta size, only in the Kit KO condition, and only when individual puncta are assessed does this modest (~10%) difference in size become statistically significant. In the revision, we eliminate this figure and focus on the per animal averages.

We interpret that the reduction in sIPSC and mIPSC frequency likely stems from a decreased proportion of functional synapse sites. The number of MLIs, their action potential generation, the density of synaptic puncta, and the ability of direct stimulation to evoke release and equivalent postsynaptic currents, are all similar in Control vs Kit KO. It is therefore feasible that a reduced frequency of postsynaptic inhibitory events is due to a reduced ability of MLI action potentials to invade the axon terminal, and/or an impaired ability for depolarization to drive (e.g. coordinated calcium flux) transmitter release. That is, while the number of MLIs and their synapses appear similar, the reduced mIPSC frequency suggests that there is a reduced proportion of, or probability that, Kit KO synapse sites that function properly.

1. Related to point 1, it would be helpful to see immunolabeling data from Kit ligand KO mice? Do these show the same pattern of reduced puncta size but no change in number?

Although we have not added a figure, we have now added experiments and a corresponding analysis in the manuscript. As we had previously for Kit KO, we now for KL KO conducted IHC for VGAT, Gephyrin, and Calbindin, and we analyzed triple-positive synaptic puncta in the molecular layer of Pcp2 Cre KL KO mice and Control (Pcp2 Cre negative, KL floxed homozygous) mice. We did not find a gross reduction in the average synaptic puncta size or density, or in the PSD-95 pinceau size. From this initial analysis, it appears that the presynaptic hypotrophy is more notable in the receptor than in the ligand knockout. We speculate that this is perhaps because the Kit receptor may have basal activity in the absence of Kit ligand, that Kit may serve a presynaptic scaffolding role that is lost in the receptor (but not the ligand) knockout, or simply that the embryonic timing of the Pax2 Cre vs Pcp2 Cre recombination events is more relevant to pinceaux development, especially as basket cells are born primarily prenatally.

1. The data using KL overexpression in PC (figure 4E,F) are intriguing, but puzzling. The reduction in sIPSC frequency and amplitude in the control PC is much greater than seen in the Kit or KL KO. The interpretation of these data, "Thus, KL-Kit levels may not set the number of MLI:PC release sites, but may instead influence the proportion of synapses that are functional for neurotransmission (Figure 4G)" is not clear and the reasoning here should be explained in more detail, perhaps in the discussion.

We have attempted to clarify this portion of the manuscript by eliminating the cartoon of the proposed model, and by revising and adding to the discussion. Either MLI Kit KO or PC KL KO seems to preserve the absolute number of MLI:PC anatomical synapse sites (IHC) but to reduce the proportion of those synapse that are contributing to neurotransmission (mIPSC). We speculate that sparse PC KL overexpression (OX) may either 1) weaken inhibition to surrounding control PCs by either diminishing KL OX PC to KL Control PC inhibition, and/or 2) act retrogradely through MLI Kit to potentiate MLI:MLI inhibition, reducing the MLI:PC inhibition at neighboring Control PCs.

Minor comments:

1. In the first sentence of the results, should "Figure 1A, B" be "Figure C, D"?

Yes, corrected.

1. The top of page 6 states "the mean mIPSC amplitude was ~10% greater in PC KL KO than in control", this does not appear to be the case in Figure 3E. control and KL KO look very similar here.

In this portion of the text citing the modest 10% increase in mIPSC amplitude, we are referring to the average amplitude of all individual mIPSC events in the PC KL KO condition; in the figure referred to by the reviewer (3E), we are instead referring to the average of all mIPSC event amplitudes per KL KO PC. Because of the dramatic difference in sample size for individual events vs cells, this modest difference rises to statistical, if not biological, significance. We include this individual event analysis only to suggest that, since we in fact saw a slightly higher event amplitude in the KL KO condition, it is unlikely that a reduced amplitude would have been a technical reason that we detected a lower event frequency.

1. Figure 3 D, duration, y-axis should be labelled "ms"

Event duration is no longer graphed or referenced. This has been replaced with total inhibitory charge.

Reviewer #2 (Recommendations For The Authors):

Methods:

Pax2-Cre line: embryonal Cre lines sometimes suffer from germline recombination. Was this evaluated, and if yes, how?

The global loss of Kit signaling is incompatible with life, as seen from perinatal lethality in other Kit Ligand or Kit mutant mouse lines or other conditional approaches. Furthermore, a loss of Kit signaling in germ cells impedes fertility. Thus, while not explicitly ruled out, since conditional Pax2 Cre mediated Kit KO animals were born, survived, and produced offspring in normal ratios, we do not suspect that germline recombination was a major issue in this specific study.

Include rationale for using different virus types in different studies (AAV vs. Lenti).

This rationale is now included and reflects the intention to achieve infection sparsity in the smaller and less dense tissue of perinatal mouse brains.

How, if at all, was blinding performed for histological and electrophysiological experiments?

It was not possible for electrophysiology to be conducted blinded for the Kit KO experiments, owing to the subjects’ hypopigmentation. However, whenever feasible, resultant microscopy images or electrophysiological data sets were analyzed by Transnetyx Animal ID, and the genotypes unmasked after analysis.

Provide justification for limiting electrophysiology recordings to lobule IV/V and why MLIs in the middle third of the molecular layer were prioritized when inhibition of PCs is dominated by large IPSCs from basket cells. Why were 2 different internals used for recording IPSCs and EPSCs in PCs and MLIs? While that choice is justified for action potential recordings, it provides poor voltage control in PC voltage clamp. Both IPSCs and EPSCs could have been isolated pharmacologically using a CsCl internal.

The rationale for regional focus has been added to the text. For MLI action potential recordings, we opted to sample the middle third of the molecular layer so that we would not be completely biased to either classic distal stellate vs proximal basket subtypes. It is our hope, in future optogenetic interrogations, to simultaneously record the dynamics of all MLI subtypes in a more unbiased way. With respect to internal solutions, we initially utilized a cesium chloride internal to maximize our ability to resolve differences in GABAA mediated currents, which was the hypothesis-driven focus of our study. While we agree that utilizing a single internal and changing the voltage clamp to arrive at per-cell analysis of Excitatory/Inhibitory input would have been most informative, our decision to utilize pharmacological methods was driven by our experience that achieving adequate voltage clamp across large Purkinje cells was often problematic, particularly in adult animals.

Introduction:

In the introduction, the authors state that inactivating Kit contributes to neurological dysfunction - their examples highlight neurological, psychiatric, and neurodevelopmental conditions.

The language has been changed.

General:

Using violin plots illustrates the data distribution better than bar graphs/SEM.

We have included violin plots throughout, and we have changed p values to numeric values, both in the interest of presenting the totality of the data more clearly.

Synapses 'onto' PCs sounds more common than 'upon' PCs.

We have changed the wording throughout.

Figure 1:

1F - there seems to be an antero-posterior gradient of Kit expression.

Though not explicitly pursued in the manuscript, it is possible that such a gradient may reflect differences in the timing of the genesis and maturation of the cerebellum along the AP axis. Regional variability is however now briefly addressed as a motivator for focused studies within lobules IV/V.

E doesn't show male/female ratios but only hypopigmentation.

This language has been corrected.

Figure 2 and associated supplementary figures:

2A/B: The frequency of sIPSCs is very high in PCs, making the detection of single events challenging. How was this accomplished? Please add strategy to the methods.

We have added methodological detail for electrophysiology analysis.

How were multi-peak events detected and analyzed? 'Duration' is not specified - do the authors refer to kinetics? If so, report rise and decay. It is likely impossible to show individual aligned sIPSCs with averages superimposed, given that sIPSCs strongly overlap. Alternatively, since no clear baseline can be determined in between events, and therefore frequency, amplitude, and kinetics quantification is near-impossible, consider plotting inhibitory charge.

Given the heterogeneity of events, we now do not refer to individual event kinetics. As suggested, we have now included an analysis of the total inhibitory charge transferred by all events during the recording epoch.

S2: Specify how density, distribution, and ML thickness were determined in methods. How many animals/cells/lobules?

For consistency with viral injections and electrophysiology, the immunohistochemical analysis was restricted to lobule IV/V. This is clearer in the revision and detail is added in the methods.

S3:

S3B: the labels of Capacitance and Input resistance are switched.

This has been corrected.

How were these parameters determined? Add to methods.

Added

In the previous figure the authors refer to 'frequency', in this figure to 'rate' - make consistent

This has been corrected.

D: example does not seem representative. Add amplitude of current pulse underneath traces.

We added new traces from nearer the group means and we now include the current trace.

F/G example traces (aligned individual events + average) are necessary.

We added example traces near the relevant group means for each condition.

Statement based on evoked IPCSs that 'synapses function normally' is a bit sweeping and can only be fully justified with paired recordings. Closer to the data would be the release probability of individual synapses is similar between control and Kit KO.

Paired recordings in both Kit Ligand and Kit receptor conditional knockout conditions is indeed an informative aim of future studies should support permit. For now, we have clarified the language to be more in line with the reviewer’s welcome suggestion.

S4:

Histological strategy cannot unambiguously distinguish MLI-PC and PC-PC synapses. Consider adding this confound to the text.

We have added this confound to the discussion.

The observation that the pinceau is decreased in size could have important implications for ephaptic coupling of MLI and PC and could be mentioned.

We agree and have added this notion to the discussion.

Y-label is missing in B.

Corrected.

Figure 3 and associated supplementary figures:

In the text, change PC-Cre to L7-Cre or Pcp2-Cre.

Changed

How do the authors explain a reduction in frequency, amplitude, and duration of sIPSCs in the KL KO but not in the Kit KO? Add to the discussion

We now address this apparent discordance in the discussion. Pax2 Cre mediates recombination weeks ahead of Pcp2 Cre. We therefore suspect that postnatal PC KL KO may be more phenotypic than embryonic MLI Kit KO because there is less time for developmental compensation. A future evaluation of the impact of postnatal Kit KO would be informative to this end.

As in Figure 2, plotting the charge might be more accurate.

We now plot total charge transfer.

Are the intrinsic properties in KL KO PCs altered? (Spontaneous firing, capacitance, input resistance).

We have added to the text that we found no difference in capacitance or input resistance between Purkinje cells from KL floxed homozygous Control animals versus those from KL floxed homozygous, PCP2 Cre positive KL KO animals. We plan to characterize both basal and MLI modulated PC firing in a future manuscript, especially since Pcp2 Cre mediated KL KO seems more phenotypic than Pax2 Cre mediated Kit KO, we agree that this seems a better testbed for investigating differences in both the basal, and the MLI-mediated modulations in, PC firing.

3D-F - Example traces would be desirable (see above, analogous to Fig. 2).

More example traces have been added.

Figure 4: 'In vivo mixtures' sounds unusual. Consider revision (e.g., 'to sparsely delete KL').

Changed

The observation that control PC sIPSC frequency is lower in KL OX PCs than in sham is interesting. This observation would be consistent with overall inhibitory synapse density being preserved. This could be evaluated with immunohistochemistry. For how far away from the injection area does this observation hold true?

Because we have now analyzed and failed to find an overt (per animal average) change in synaptic puncta size or density in the whole animal Control vs PCP2 Cre mediated KL KO conditions, we do not have confidence that it is feasible to pursue this IHC strategy in the sparse viral-mediated KL KO or OX conditions. To the reviewer’s valid point however, we intend to probe the spatial extent/specificity of the sparse phenomenon when we are resourced to complement the KL/Kit manipulations with transgenic methods for evaluating MLI-PC synapses specifically, potentially by GRASP or related methods that would not be confounded by PC-PC synapses. Transgenic MLI access would also facilitate determining the spatial extent to which opto-genetically activated MLIs evoke equivalent responses in Control vs KL manipulated PCs.

Y-legend in D clipped.

Corrected

Existing literature suggests that MLI inhibition regulates the regularity of PC firing - this could be tested in Kit and KL mutants.

For now, based upon transgenic animal availability, we have now included an evaluation of PC firing in the (Pax2 Cre mediated) Kit KO condition. PC average firing frequency, mean ISI, and ISI CV2 were not significantly different across genotypes. A KS test of individual ISI durations for Control vs Kit KO did reveal a difference (p<0.0001). We have added a supplementary figure (S6) with this data. It is possible that in the more phenotypic PC KL KO condition that we may find a difference in these PC spiking patterns of PC firing, however, we are also eager to test in future studies whether postnatal KL or Kit KO impairs the ability of MLI activation to produce pauses or other alterations in PC firing or in PF-PC mediated plasticity.

Reviewer #3 (Recommendations For The Authors):

Reference to Figure 1A in the Results section is slightly inaccurate. Kit gene modifications are illustrated in Figures 1A, B. Where Figure 1A shows Kit distribution. Please rephrase. Relatedly, the reference to Figs 1B - D are shifted in the results section, and 1E is skipped.

We have changed the text.

Please show cumulative histograms for frequency too for consistency with amplitude (e.g. Fig 2).

We have instead, for reasons outlined by other reviewers, documented total charge transfer for both Kit KO and KL KO experiments where sIPSC events were analyzed.

Fig S3: include example traces of PPR.

This is now included.

Include quantifications of GABAergic synapse density in Fig S4.

This is now included.

Include inset examples of KO in Fig S4A.

This is now included.

Add average puncta size graphs along Figure S4B. The effect apparent in the histogram of S4B is small and statistics using individual puncta as n values (in the 20,000s) therefore misleading.

Per animal analysis is now instead included in the figure and text.

Figure S4B y axis label blocked.

Corrected

Include quantification referenced in "As PSD95 immunoreactivity faithfully follows multiple markers of pinceaux size 40, we quantified PSD95 immunoreactive pinceau area and determined that pinceaux area was decreased by ~50% in Kit KO (n 26 Control vs 43 Kit KO, p<0.0001, two-tailed t-test)."

We added a graph of per animal averages, instead of in text individual pinceau areas.

Include antibody dilutions in the methods.

Added.

It's unclear from the text where the Mirow lab code comes from.

Detail has now been added in text.

Typo in methods "The Kit tm1c alle was bred...".

Corrected

Typo in Figure S4 legend "POSD-95 immuno-reactivity".

Corrected

Version 2

Abstract

Summary

The cell-type specific expression of ligand/receptor and cell-adhesion molecules is a fundamental mechanism through which neurons regulate connectivity. Here we determine a functional relevance of the long-established mutually exclusive expression of the receptor tyrosine kinase Kit and the trans-membrane protein Kit Ligand by discrete populations of neurons in the mammalian brain. Kit is enriched in molecular layer interneurons (MLIs) of the cerebellar cortex (i.e., stellate and basket cells), while cerebellar Kit Ligand is selectively expressed by a target of their inhibition, Purkinje cells (PCs). By in vivo genetic manipulation spanning embryonic development through adulthood, we demonstrate that PC Kit Ligand and MLI Kit are required for, and capable of driving changes in, inhibition of PCs. Collectively, these works in mice demonstrate that the Kit Ligand/Kit receptor dyad sustains mammalian central synapse function and suggest a rationale for the affiliation of Kit mutation with neurodevelopmental disorders.

Introduction

The proto-oncogene receptor tyrosine kinase Kit (c-Kit, CD117) is an evolutionarily conserved, loss-of-function intolerant, gene that is enriched in specific neuron populations and tentatively associated with rare cases of neurological dysfunction (pLI= 0.98 and LOEUF of 0.17) (Figure S1A,B) [1, 2]. It is perhaps unsurprising that loss-of-function Kit mutations in humans are rare; Kit is highly pleiotropic, supporting diverse cell populations including melanocytes, hematopoietic stem/progenitor cells, germ cells, masts cell, and interstitial cells of Cajal [3]. Thus, Kit mutations can result in diverse conditions such as hypo- or hyperpigmentation or melanoma, anemia or leukemia, infertility, mastocytosis, and impaired gut motility or gastro-intestinal tumors (Kit biology reviewed [2] [4]).

The activation of Kit is stimulated by Kit Ligand (Gene symbol KITLG, herein KL), a single-pass transmembrane protein having an extracellular active domain that induces Kit dimerization and kinase activity. Across several decades and species, it has been reported that KL and Kit are mutually expressed in discrete neuron populations known to be connected; it has thus been hypothesized that the expression pattern of KL-Kit may reflect a role in connectivity [5-8]. Consistent with such a function, case reports have implicated inactivating Kit mutations with disorders of the central nervous system (References [9-13] and Figure S1C). Clinical phenotypes include developmental delay, ataxia, hypotonia, intellectual disability, deafness, and autism spectrum disorder. White matter abnormalities have been noted, and limited evidence suggests KL-Kit may indeed regulate axon outgrowth. KL or Kit mutation or pharmacological manipulation alters outgrowth of spinal commissural axons in vitro [14], KL stimulates cortical neurite outgrowth in vitro [15], and Kit reduction in cortex of developing rats or mice delays axon extension [16]. While these reports provide anatomical/morphological data suggesting that KL-Kit influences connectivity, these studies did not address consequences to neuronal physiology or synaptic connectivity. Functional studies of KL-Kit in synaptic physiology have likely been hampered by pleiotropy and organismal or cell death. Severe global KL-Kit hypomorphs exhibit fatal anemia, and developmental Kit depletion or inhibition in the cerebrum has resulted in the death of neuronal progenitors or nascent neurons [16-18]. Therefore, whether KL-Kit is necessary for synaptic function has remained largely unstudied. Here, we address this gap.

In mice, rats, and humans, Kit is abundantly expressed by molecular layer interneurons (MLIs) of the cerebellar cortex [6-8, 19], while cerebellar KL is restricted to Purkinje cells (PCs), which MLIs provide GABAergic synaptic inhibition to. We tested the hypothesis that PC KL and MLI Kit were essential to the GABAergic inhibition of PCs. We created a Kit conditional knockout mouse and accomplished embryonic knockout of Kit from MLIs; separately, we knocked out KL from postnatal PCs. By either method, disruption of the KL-Kit pair produced robust and specific impairments to GABAergic inhibition of PCs. Through sparse postnatal viral manipulation of KL in PCs, we provide evidence that KL functions throughout adulthood to regulate synaptic function. These results demonstrate that cell-type specific expression of KL and Kit influences functional connectivity in the mammalian brain, informing the long-observed expression of KL-Kit in the brain and suggesting a rationale for the affiliation of Kit loss with neurological impairment.

Methods

Animals

All procedures performed at Michigan State University were approved by the Institutional Animal Care and Use Committees of Michigan State University, which is accredited by the Association for Assessment and Accreditation of Laboratory Animal Care (AAALAC). Mice were on a 12-hour dark/light cycle, with ad libitum food and water. Male and female mice were used for all experiments. Mice were genotyped in-house using standard PCR-based methods, or via Transnetyx.

Creation of Kit conditional knockout mouse

We generated a strain of mice having exon 4 of Kit flanked by LoxP sites (“floxed”). This was accomplished by utilizing the services of UNC Animal Models Core (Dr. Dale Cowley) to generate chimeras from an ES cell line having a Knockout-first Allele (with conditional potential) for Kit (HEPD0509_8). The ES clones were obtained through the EuMMCR (European Mouse Mutant Cell Repository) of the EUCOMM (European Conditional Mouse Mutagenesis Program) [20]. Chimeric mice were screened for germline transmission of the targeted allele and were then crossed to a Flp deleter strain (C57BL/6N-Albino-Rosa26-FlpO) to convert the knockout-first allele to a conditional allele (tm1c) through removal of the FRT-flanked selection cassette. F1 male offspring carrying Kit tm1c and Flp alleles were crossed to C57Bl6/J females; we selected resultant progeny heterozygous for Kit tm1c but negative for Flp and the selection cassette. The Kit tm1c allele was bred to homozygosity while selecting against animals carrying the Tyr c-Brd allele (previously conveyed by C57BL/6N-Albino-Rosa26-FlpO). The resultant strain of mice having the Kit floxed tm1c allele are given the nomenclature C57BL/6N−Kittm1c(EUCOMM)Mirow/J; the Mirow lab code is registered through the National Academies to the author and principal investigator herein, Michael Roland Williams. Additional strains of mice used were Pax2-Cre transgenic mice (STOCK Tg(Pax2-cre)1Akg/Mmnc, RRID:MMRRC_010569-UNC [21], Kit eGFP (Tg(Kit-EGFP)IF44Gsat/Mmucd; Gensat) [22] backcrossed for greater than 5 generations to C57Bl6/J (Stock No: 000664, The Jackson Laboratory), KitL floxed (Kitltm2.1Sjm/J, Stock #017861) [23], and Pcp2-Cre (Stock # 004146) mice [24], via The Jackson Laboratory.

To produce mixed litters of “Control” and “Kit KO” animals, mice homozygous for the Kit tm1c allele but negative for Pax2-Cre were crossed to mice homozygous for the floxed allele and Pax2-Cre positive. We used Kit KO females crossed to Control males as our breeding scheme. Experimental animals were derived from at least 5 generations of backcrossing Kit KO animals to Kit tm1c homozygous animals derived from a separate (Kit tm1c homozygous x Kit tm1c homozygous) breeding cohort.

Validation of Kit knockout

We investigated conditional knockout of Kit protein from the cerebellum by Western blot of total protein lysates from acutely harvested cerebellums (∼P52) and by indirect immunofluorescence on free-floating vibratome-generated sections of transcardial-perfused formaldehyde-fixed mouse brains (∼P31). Western blots were probed with Rabbit: α-Kit (1:1000, clone D13A2, Cell Signaling Technologies product 3074), α-GAPDH (1:2000, Clone 14C10, product 2118 Cell Signaling Technology), or α-parvalbumin antibody (1:3000, product PV-27, Swant), each primary was detected via HRP conjugated Goat α-Rabbit secondary antibody (1:4000, Bio-Rad, product 170-6515) after incubation with Clarity ECL substrate (Bio-Rad, 1705061), via a Bio-Rad ChemiDoc MP.

For subsequent imaging and electrophysiology, we focused on lobule IV/V of the cerebellar cortex, as this area was superficially accessible for surgery, and provided a consistent area for focused comparisons between animals that would reduce the influence of spatial variations in KL or Kit expression.

For indirect immunofluorescence, primary antibodies were rat α-Kit (1:500, Rat α-Kit clone ACK4, ThermoFisher, product MA5-17836), mouse α-calbindin (1:2000, α-calbindin-D28K, CB-955, Sigma Aldrich), and rabbit α-parvalbumin (1:500, PV-27, Swant). Primaries were detected by Goat-host highly cross-adsorbed secondary antibodies coupled to Alexa 488, Cy3, or Alexa 647 (1:400, Jackson ImmunoResearch). Nuclei were labeled with DAPI, tissues were mounted in an anti-fade reagent, and fluorescent signals captured laser scanning confocal microscopy. Cell densities for MLIs (parvalbumin positive and/or calbindin negative somas of the molecular layer) were performed on Maximum Z-projections of 2.5 to 2.6 micron thick stacks, using the multi-point feature for counting somas, and the polygon ROI tool for determining the area of the Molecular Layer in which the somas were counted; results from two-technical replicates were averaged for each animal. The polygon tool was also used for determining maximal PSD-95+ pinceau area. To determine the width of the molecular layer, seven linear ROIs per field of view of comparable locations of lobules IV/V were averaged to generate per animal layer thickness measurements. For analysis of synapses, we quantified puncta (of 0.05-5 μm2 using the Analyze Particles function) in the molecular layer that were triple positive (each channel auto-threshold) for rabbit α-VGAT(1:500, 131 003), mouse α-Gephyrin (1:5000, 147 011), and Chicken α-Calbindin (1:500, 214 006), all from Synaptic Systems. For analysis of pinceau markers we utilized mouse α-Kv1.1 (1:250, K36/15), Kv1.2 (1:250, K14/16), and PSD-95 (1:200, K28/43), all from NeuroMab/AntibodiesInc.

Stereotaxic Procedures

Injections of replication-defective viral particles used protocols substantially similar to those previously published [25]. In brief, mice were brought to an anesthetic plane using inhaled isoflurane by a low-flow system (SomnoSuite, Kent); thermal regulation was assisted by heating pad, and pain managed through topical lidocaine and by intraperitoneal ketoprofen. A digital stereotaxic device (Kopf) and reference atlas were used to localize a small hole in the skull; cerebellar coordinates were: neonates; A/P, -2.55; M/L, - 1.1 (reference Lambda), D/V, 1.5 to 0.5 from dura; juvenile; A/P, -6.25; M/L, - 1.35; D/V, 1.5 to 0.75 (reference Bregma); adults; A/P, -6.35; M/L, - 1.8; D/V, 1.5 to 0.75 (reference Bregma). Viral particles were delivered via a ∼30-gauge Hamilton Syringe controlled by digital syringe pump (WPI). The infused virus volume was 1-2 ul and there was a 5-minute period after infusion prior to needle withdrawal. Animals were monitored during and after operations. Ages for injection were, nominally, postnatal day (P) 7, 19, and 56; for P7 and P19, variability of up to two days was utilized to accommodate differences in animal size, and for P56 the eighth week of life was utilized.

Viruses

Lentiviral particles were derived from our vectors previously described to express transgenes under the hUbiC promoter and pseudotyped with VSV-G [26, 27]. Silent mutations disrupted the internal EcoRI sites of mouse KitL (MC204279, OriGene), the KitL coding sequence was then flanked by EcoRI sites, and the subsequent KitL coding sequence was introduced into a FUC-T2A-(EcoRI) plasmid, to create a FUC-T2A-KL-1 plasmid, wherein C designates mCherry. Subsequent deletion of KitL Exon 6 generated a KL-2 vector [28-31]. Sequences were verified by sequencing. Adeno-Associated Viral (AAV) particles were created by Vigene to drive Cre expression by the Purkinje Cell-Specific L7.6 promoter [32], or to drive mCherry or mCherry-T2A-KL-2 in a Cre-On fashion under the EF1α promoter, with the AAV1 serotype.

Electrophysiology

For patch clamp electrophysiology of Control vs Kit KO, cells were recorded in slices generated from animals at 36.9 ± 0.6 and 37.4 ± 0.9 days for MLIs and PCs respectively. Analysis of the Pcp2 Cre mediated KL KO tissues was conducted similarly at ∼40-44 days old. For viral PC KL KO animals injected as neonates (P7), juvenile (P19), and adults (P56) were recorded at an average of ∼35, 45, and 83 days old, respectively. Avertin-anesthetized mice were perfused with a carbogen-equilibrated ice-cold slicing solution containing (in mM): 110 C5H14ClNO, 7 MgCl2.6H2O, 2.5 KCl, 1.25 NaH2PO4, 25 NaHCO3, 0.5 CaCl2-2H2O, 10 Glucose, and 1.3 Na-Ascorbate. In the same solution, we generated 250-micron thick parasagittal sections of the cerebellum (Leica VT1200). Slices recovered at 34°C for 30 minutes in a carbogenated ACSF containing (in mM): 125 NaCl, 25 NaHCO3, 1.25 NaH2PO4, 2.5 KCl, 1 MgCl2.6H2O, 1 CaCl2-2H2O and 25 Glucose; slices were then held at room temperature for at least 30 minutes before recording. Recordings were performed in carbogenated external recording ACSF (32.7 ± 0.1°C) containing (in mM): 125 NaCl, 25 NaHCO3, 1.25 NaH2PO4, 2.5 KCl, 1 MgCl2.6H2O, 2 CaCl2-2H2O and 25 Glucose. MLIs and PCs of folia IV/V were targeted for recording by IR-DIC, or epifluorescence stimulated by a CoolLED pE-4000, detected via a SciCam Pro on a SliceScope Pro 6000-based rig (Scientifica). Recording electrodes were pulled (Narishige, PC-100) from standard-wall borosilicate glass capillary tubing (G150F-4, Warner Instruments) and had 5.1 ± 0.08 and 2.8 ± 0.02 MΩ tip resistance for MLIs and PCs respectively. Spontaneous Inhibitory Postsynaptic Currents (sIPSCs) and Miniature Inhibitory Postsynaptic Currents (mIPSC) were recorded with an intracellular solution containing (in mM):140 CsCl, 4 NaCl, 0.5 CaCl2-2H2O, 10 HEPES, 5 EGTA, 2 Mg-ATP, and 0.4 Na-GTP, 2 QX-314. For action potential recordings and for Miniature Excitatory Postsynaptic Currents (mEPSCs), the intracellular solution contained (in mM): 140 K-gluconate, 10 KCl, 1 MgCl2, 10 HEPES, 0.02 EGTA, 3 Mg-ATP, and 0.5 Na-GTP as described previously [33]. The internal pipette solution pH was adjusted to 7.35 with CsOH for IPSCs and mIPSCs, and with KOH for mEPSCs and action potentials, while for all the osmolarity was adjusted to 300 mOsmolL-1 with sucrose. In whole-cell voltage clamp mode PCs were held at -70 mV; to isolate sIPSCs, CNQX or NBQX (10 μM) and D-AP5 (50 μM) were added to the recording solution and mIPSCs recordings additionally included 1 μM tetrodotoxin (TTX). To record mEPSCs, 1 μM TTX and 50 μM picrotoxin was added to extracellular solution. Action potentials were recorded in MLIs within the middle third of the molecular layer; spontaneous action potentials were recorded in I=0 mode, or action potentials were evoked by depolarizing current injection in 600 ms steps of 10 pA, from 0 to 150 pA, with a 7 second inter-sweep interval. For paired pulse, current ranging from 150 uA to 250 uA was delivered by concentric bipolar stimulating electrode in the molecular layer. Stimulus duration was 50 μS, and ISI was 50 ms. For each cell, 10 sweeps were recorded with 10 S inter-sweep interval. A square-wave voltage stimulation pulse was utilized to determine input resistance and cell capacitance. Purkinje Cells with an access resistance of 10-20 (15.9 ± 0.5), or MLIs with an access resistance <30 (23.6 ± 0.6) MΩ were considered for recording. Recordings with >20% change in series resistance were excluded from analysis. Signals were acquired at 10 KHz with a low-noise data acquisition system (Digidata 1550B) and a Multiclamp700-A amplifier and were analyzed using pClamp11.1 (Molecular Devices) after low pass Bessel (8-pole) filtration (3dB cut off, filter 1 KHz). The minimum amplitude threshold for detecting IPSCs and EPSCs was 15 pA; for mIPSCs and mEPSCs the cutoff was 8 pA. For determining frequency and amplitude, individual events longer than 1 ms were included while overlapping events were manually rejected from analysis. Inhibitory charge is reported as the per-cell sum over the recording epoch of the area under the curve for all pharmacologically resolved inhibitory events. For the analysis of minis, both mIPSCs and mEPSCs with over 8 pA amplitude and longer than 1 ms duration were included while overlapping events were rejected from analysis; also rejected from analysis was any miniature event for which amplitude, rise, and decay metrics were not all available.

Software

Microscopic image processing was conducted via FIJI/ImageJ, electrophysiological data was analyzed by pClamp11.1, statistical analyses were conducted via GraphPad Prism 9, and Figures were created with BioRender, PowerPoint, and GraphPad Prism 9.

Results

We generated mice in which Kit exon 4 is flanked by LoxP sites (Kit tm1c); Kit gene modifications are illustrated in Figure 1C, D. Kit tm1c homozygous Control animals had normal appearance, sex ratios, and reproductive success. Kit is abundantly expressed in parvalbumin positive GABAergic interneurons in the molecular layer of the cerebellar cortex (MLIs), whereas Kit Ligand (KL) is expressed by Purkinje cells (PCs) (Figure 1A, B). To determine if Kit influences synapse function, we conditionally knocked out Kit from MLIs and assessed the MLI:PC synapse. Pax2 is a transcription factor expressed early in the lineage of GABAergic interneurons of the cerebellum [34]; a Pax2-Cre transgene enables embryonic recombination [35, 36]. We generated Kit tm1c homozygous litters of which nominally half were hemizygous for Pax2 Cre (Kit KO) and half were not (Control). We produced Control and Kit KO animals in normal sex and genotype ratios. As expected for the C57Bl6/J background, Control littermates had dark coat and eyes, but littermate Kit KO animals had white whiskers, white fur with variable pigmented patches, and black eyes; example littermates Figure 1E. Given Pax2 expression in melanocytes [37] and the role of Kit in melanogenesis (Reviewed in [38, 39]), the pigmentation phenotype provided gross confirmation of conditional Kit KO. We thus confirmed that cerebella of Kit KO animals were Kit depleted. By immunohistochemistry in Control animals, we recapitulated the known pattern of Kit immunoreactivity in the parvalbumin positive MLIs of the cerebellar cortex (Control, 1F and Inset). In contrast, a sex-matched Kit KO littermate demonstrated loss of cerebellar Kit immunoreactivity (Kit KO, 1F and Inset). In data not shown, we confirmed that both mGlur1/2 and neurogranin positive cerebellar Golgi cells (which share the MLI Pax2+ lineage, and of which some are transiently Kit positive) were still present in the Kit KO condition [34, 40-42]. By Western blot of cerebellar lysates from Control animals, Kit resolved as a major band at ∼120 kD, but in Kit KO littermates, Kit immunoreactivity was lost; GAPDH as loading control was equivalent, as was parvalbumin (which marks mature MLIs and PCs) Figure 1G.

Design and validation of a Kit conditional knockout mouse.

(A, B) Kit receptor tyrosine kinase (Kit) is enriched in parvalbumin positive GABAergic interneurons of the molecular layer (i.e. basket and stellate cells, MLIs) of the cerebellar cortex, where they synapse onto each other and onto Purkinje cells (PCs, Calbindin+), which express Kit Ligand (KL). Scale bar 10 microns. Expression pattern schematized in B, Control.

(C)In humans and mice, Kit is encoded by up to 21 exons, which in mouse is encoded on the plus strand of chromosome 5 at 75,735,647-75,817,382 bp. We generated a Kit conditional knockout mouse in which Kit Exon 4 is floxed, flanked by LoxP sites.

(D)We generated Control mice homozygous for the Kit floxed allele Kittm1c(EUCOMM)Mirow, which varied in Pax2-Cre)1Akg transgene status, with the goal of depleting Kit from MLIs in embryonic development.

(E)Pax2 Cre mediated Kit KO mice were notably hypopigmented in hair and whiskers, though not eyes.

(F)Confocal microscopy of Kit immunoreactivity in cerebella from age and sex matched Control and Kit KO littermates demonstrates the established enrichment of Kit in the molecular layer of the Control cerebellar cortex, and its loss in Kit KO. Scale Bar 500 microns top row, 50 microns inset.

(G)Utilizing a distinct assay and different primary antibody, we confirm the detection of Kit immunoreactivity in Controls, and its loss in Kit KO litter mates of either sex by Western Blot of total protein lysates of cerebella. We affirmed equivalent protein loading by GAPDH and parvalbumin.

Having validated that Kit KO animals indeed lacked cerebellar Kit, we then assessed the functional impact. We compared spontaneous GABAergic inhibitory post synaptic currents (sIPSC) in PCs of acute slices generated from ∼P34 Control and Kit KO animals, schema Figure 2A, example traces Figure 2B. Kit KO was associated with a significant change in the distribution of sIPSC event amplitudes (Figure 2C). For each PC of Control or Kit KO conditions, we calculated the average sIPSC event frequency and amplitude, as well as total inhibitory charge transfer over the recording epoch. We determined there was a significant decrease in sIPSC frequency (Figure 2D) and a less severe decrease of sIPSC amplitude and inhibitory charge. This reduced inhibition did not produce obvious impacts to spontaneous PC firing dynamics: we determined (in 29 Control vs 22 Kit KO cells) that PC membrane potential and membrane resistivity were not significantly different, and that per cell average: spontaneous action potential frequency, inter-spike interval (ISI), and ISI CV2 were not significantly different, despite some difference in the distribution of individual ISI across genotypes.(Figure S6). Though circuit consequences of the Kit KO mediated PC disinhibition were thus not immediately apparent, the reduced sIPSC frequency in Kit KO was not associated with changes in the density or distribution of MLIs or in the thickness of the molecular layer (Figure S2).

The knockout of Kit from Cerebellar Cortex Interneurons Impairs Inhibition of Purkinje Cells. 419

(A, B) Experimental schema and example traces of spontaneous inhibitory postsynaptic currents in Purkinje Cells (PCs) from Control animals or from those with Kit KO from cerebellar cortex molecular layer interneurons (MLIs). Scale bar is 500 ms x 100 pA.

(C)A frequency distribution plot for individual sIPSC event amplitudes recorded in PCs as in A,B. A KS test reveals a significant difference in the distribution of these event amplitudes; p<0.0001, n in chart.

(D)For each PC, the average sIPSC Frequency and Amplitude, and the total Inhibitory Charge transfer, was determined. There was a significant ∼50% decrease in sIPSC frequency, but the decrease in Amplitude or Charge transfer was not significant.

(E-G) Experimental schema: In separate experiments, we recorded miniature postsynaptic currents from PCs or from MLIs in Control and in Kit KO animals.

(H) Example traces of mIPSCs or (I) mEPSCs recorded in PCs, and example traces of mIPSC in MLIs, all from Control or from Kit KO animals. Scale for H and J is 50 pA, 25 pA for I, all 500 ms.

(K-M) Analysis of per cell average miniature event Frequency and Amplitude revealed that Kit KO significantly reduced mIPSC frequency, but not amplitude, in PCs by >50%, p=0.013. mEPSCs in PCs and mIPSC in MLIs were not significantly different between Control and Kit KO.

n in charts, refers to the number of cells. Error bars are SEM. p-values were calculated by a two-tailed t-test, with Welch’s correction as needed.

Since the number and distribution of MLIs were normal in Kit KO, we sought to determine if the reduced PC inhibition was related to altered MLI physiology (Schema, Figure S3A). However, capacitance, input resistance, and spontaneous or evoked MLI action potential frequency was not different between Control and Kit KO (Figure S3B-E). Since MLIs numbers, distribution, and firing were apparently normal in Kit KO, we sought to determine if there were defects in evoked neurotransmitter release. Direct electrode stimulation of the outer or inner molecular layer in a paired-pulse paradigm (Schema, Figure S3F,J; example traces Figure G,K) evoked IPSCs of equivalent mean amplitude in PCs from Control and Kit KO (Figure S3H,L), and we detected no difference in the paired pulse ratio between conditions (Figure S3I,M). These data suggest that direct stimulation can evoke a similar maximum of release probability between Control and Kit KO conditions. We therefore investigated markers of the MLI:PC synapses to determine if they were altered.

The density and average size of GABAergic synapses onto PCs (triple positive VGAT, Gephyrin, and Calbindin puncta in the molecular layer) was not reduced by Kit KO (example immunofluorescence, Figure S4A, Quantification S4B,C). MLI axon terminals contribute not just to individual axo-dendritic synaptic puncta, but also to interdigitated MLI axon collaterals around the soma and initial axon segment of PCs, forming “pinceaux” structures. By parvalbumin immunoreactivity, these structures were preserved but smaller in Kit KO (arrowheads Figure S4C), and this was confirmed by decreased Kv1.2 (Figure S4D) and PSD-95 (Figure S4E). As PSD-95 immunoreactivity faithfully follows multiple markers of pinceaux size [43], we quantified PSD-95 immunoreactive pinceau area and we determined that pinceaux size was decreased by ∼50% in Kit KO (26 Control vs 43 Kit KO pinceau from n of 5 Control vs 8 Kit KO animals, Figure S4F). These results suggested that Kit KO MLIs may have defects in physical and/or functional maturation of synapses onto PCs.

To evaluate the number of functional GABAergic synapses, we analyzed miniature synaptic currents in acute cerebellar slices, (schema Figure 2 E, F, G; example traces Figure 2 H, I, J). In PCs of Kit KO, there was a significant ∼50% reduction in the average frequency, but not amplitude, of mIPSCs (Figure 2K). In contrast, the average frequency and amplitude of mEPSCs in PCs from Control and Kit KO did not differ (Figure 2L). These data suggested that Kit KO impaired the MLI:PC synapse but did not rule out a general defect in MLI synapse function. As MLIs synapse not just onto PCs, but also onto other MLIs, we evaluated mIPSCs in MLIs. Mean mIPSC frequency and amplitude in MLIs was unchanged between Control and Kit KO (Figure 2M). Therefore, embryonic MLI Kit KO was associated with a specific defect in the number or proportion of functional GABAergic synapses onto PCs. As MLI Kit and PC KL expression is maintained postnatally, we hypothesized that postnatal postsynaptic PC KL KO would phenocopy embryonic presynaptic MLI Kit KO.

We utilized previously described KL floxed [23] and Pcp2-Cre [24] strains to yield mice with Control or KL KO PCs (Schema Figure 3A). We confirmed depletion of cerebellar KL transcripts by rtPCR in data not shown[23], and we confirmed that PC KL KO did not grossly disrupt the pattern of Kit immunoreactivity. Patch clamp of PCs in acute slices revealed that PC KL KO produced significant differences in the distribution of sIPSC event amplitudes (Example traces Figure 3B, Figure 3C). PC KL KO reduced the average sIPSC frequency and amplitude, as well as the total inhibitory charge transfer (“Charge”) recorded in PCs (Figure 3D), without changes to PC capacitance or membrane resistance. The phenotype was specific to GABAergic synapses: there was a substantial reduction in the frequency but not the amplitude of mIPSC events in PCs (Figure 3E). As with Kit KO, we detected no decrease in the per-animal average size or density of triple positive VGAT/Gephyrin/Calbindin synaptic puncta in the molecular layer (n=4 Control vs 4 KL KO animals, mean area 0.249 vs 0.286 μm2, p=.1962; mean density 0.045 vs 0.046 per μm2, p=.9658), Not Illustrated). We further detected no difference either in PDS-95+ pinceau area (49 Control vs 48 KL KO pinceaux from n=4 Control vs 4 KL KO animals, mean area 57.93 vs 53.37 μm2 p=.6036, Not Illustrated). Thus, as with Kit embryonic Kit KO, postnatal KL KO apparently reduced the proportion of functional, if not the absolute structural total, of GABAergic synapse sites upon PCs. In contrast to this impact on GABAergic synapses, the frequency and amplitude of mEPSCs in KL KO PCs was not different vs Controls (Figure 3F). In data not shown, a KS test revealed a significant difference in the individual mEPSC event amplitude distribution (p=0.0002, n=1196 Control vs 1244 KL KO), however the mean mEPSC amplitude was within 1% between conditions. Similarly, a KS test of mIPSC event amplitudes revealed that the mean mIPSC amplitude was ∼10% greater in PC KL KO than in Control (p<0.0001, n=5996 Control vs 3958 KL KO). It is therefore unlikely that decreased event amplitude contributed to the specific and robust >50% decrease in mIPSC frequency detected in PC KL KO. Thus, as with MLI Kit KO, PC KL KO specifically reduced synaptic inhibition of PCs.

Knockout of Kit Ligand from Purkinje Cells decreases the inhibitory input they receive.

(A, B) Experimental schema and example traces of inhibitory postsynaptic currents detected in Purkinje Cells (PCs) from Control animals or from those with Kit Ligand knockout (KL KO) accomplished by a Pcp2-Cre x KL floxed strategy. Scale bar is 500 ms x 100 pA.

C) A frequency distribution plot for individual sIPSC event amplitudes recorded in PCs as in A, B. A KS test reveals a significant difference in the distribution of these event amplitudes; p<0.0001, n in chart.

(D) For each PC, the average sIPSC Frequency and Amplitude, and the total inhibitory Charge transfer, was determined. There was a significant ∼40% decrease in sIPSC Frequency, Amplitude, and Charge transferred to KL KO PCs.

(E, F) The average mIPSC recorded in PC KL KO vs Control PCs had a >50% decreased Frequency p<0.0001 without a corresponding decrease in Amplitude. The average mEPSC Frequency and Amplitude recorded in separate PCs did not differ between Control and KL KO. Scale bar is 500 ms x 100 pA for mIPSC, 500 ms x 50 pA for mEPSC.

n in charts refers to number of cells. Error bars are SEM. p-values were calculated by a two-tailed t-test, with Welch’s correction as needed.

Recombination by Pcp2 Cre begins at ∼P7 [24], whereas Pax2 Cre recombination occurs at ∼E15 [21, 34]. Our results thus suggest that continued expression of the KL-Kit dyad sustains inhibitory drive to PCs. To test this hypothesis (and to avoid transformative effects of Kit overexpression), we tuned KL postnatal expression in vivo.

To determine if acute focal manipulations in KL modulate PC inhibition, we injected replication-defective viruses leveraging the PC-specific L7.6 promoter [32] to express Cre in KL floxed animals, and we identified Cre-positive KL KO PCs by mCherry expression (Schema Figure 4A, Example transduction, Figure S5A-C,E-G). Injections were at P7, P18, or P56; acute slices were generated 2-3 weeks after injections. In the P7 neonatal animals we utilized lentivirus owing to its reduced spread vs AAV particles, which were used for P18 and P56. Patch-clamp recordings in areas of sparse infection revealed that regardless of when KL was depleted, PC KL KO reduced PC sIPSC frequency and inhibitory charge transferred (P7, P18, and P56 in Figure S5 D, H, I; P18 injections Figure 4 B, C, D). We next examined the effects of PC KL over-expression using a complementary strategy. We co-injected L7.6 Cre AAV with AAV encoding EF1α driven Cre-on mCherry and the membrane bound isoform of mouse KL (mKL2) via a T2A element to produce PC KL overexpression (OX) (Schema Figure 4E). Whether at P18 (shown) or P56 (not shown), KL OX PCs had increased sIPSC frequency, amplitude, and inhibitory charge transferred, compared to nearby Control PCs (Figure 4F, G, H). We noted that Control PCs from the PC KL KO paradigm received a similar degree of inhibition (frequency, amplitude, charge) as did PCs from Sham injections. In contrast, while the KL OX PCs received elevated GABAergic input compared to their adjacent Control PCs, the (KL OX) relative local increase in inhibitory drive was not beyond the level of inhibition recorded in PCs from separate Sham animals (Figure 4F, G, H). That is, the local gain in synaptic input to KL OX PCs appears to come at the expense of decreasing inhibition to nearby Control PCs.

Local levels of Kit Ligand influence the inhibition Purkinje cells receive.

(A) To accomplish in vivo Control and sparse Kit Ligand Knockout (KL KO) Purkinje Cells (PCs), an AAV encoding Cre under the PC specific L7.6 promoter was co-injected with an AAV encoding a Cre-On mCherry cassette under the Ef1α promoter.

(B, C, D) In animals injected at P18, PC KL KO neurons demonstrated a ∼70% decrease in sIPSC event frequency compared to adjacent Control PCs (p<0.0001) or PCs recorded from Sham control animals (p=0.0002). Control vs Sham sIPSC frequency was not significantly different. The same pattern was found for both mean sIPSC Amplitude (C) and for total inhibitory Charge transferred (D).

(E)To accomplish in vivo Control and sparse Kit Ligand overexpressing (KL OX) PCs, the L7.6 Cre AAV was co-injected with AAV expressing mCherry and (T2A) murine Kit Ligand isoform 2 under the Ef1α promoter.

(F, G, H) In animals injected at P18, PC KL OX neurons demonstrated an ∼8-fold increase in sIPSC Frequency vs neighboring Control PCs (p<0.0001). The frequency of sIPSC events in KL OX PCs was not significantly different from PCs recorded in different Sham animals; this Sham PC sIPSC frequency (while comparable across studies) was 7-fold higher than Control PCs within the KL OX experimental animals (p<0.0001). A similar pattern was found for sIPSC amplitude (G) and for total inhibitory Charge transferred (H).

n in charts, refers to number of cells. Error bars are SEM. p-values were calculated by Brown-Forsythe ANOVA test with Dunnett’s T3 multiple comparisons test.

Discussion

For genes essential to survival or reproduction, germline knockout strategies are not feasible for evaluating postnatal physiology. Among such genes, Kit is an example of a highly pleiotropic gene, and so the interpretation of global hypomorphs is challenging and the recombination of conditional alleles outside of desired tissues can lead to lethal or sterile phenotypes. By generating a conditional knockout mouse, we were able to produce viable animals with Kit depleted from the cerebellum to reveal a role in synapse function.

MLI number, distribution, action potential firing, and the density of GABAergic synaptic puncta onto PCs, are all apparently normal in MLI Kit KO. These data suggest that while MLI axon terminals are present in Kit KO, they may be less functional. An increased failure rate between action potential generation and neurotransmitter release would be consistent with the normal firing frequency and synapse puncta density but decreased sIPSC and mIPSC frequency observed. Under the PPR paradigm, the apparently equivalent evoked responses may be due to direct stimulation of axon terminals and/or simultaneous recruitment of multiple MLIs.

Kit KO most directly impacts the MLI:PC synapse, based on seemingly normal miniature events for parallel fiber: PC synapses (mEPSC in PC) and MLI:MLI (mIPSC in MLI) synapses. This is perhaps expected given the cell type specific expression of KL by PCs and Kit by MLIs, however, it is a critical distinction as Cerebellins [44] and Calsystenin [45] impact multiple forms of synaptic input to PCs. It will be informative to determine if the residual inhibitory drive to PCs observed in MLI Kit KO or PC KL KO comes from PC axon collaterals [46-49] whose input we might expect to be preserved if KL-Kit functions primarily to mediate connectivity between different cell types, or perhaps from a subtype of interneuron not dependent upon Kit signaling. This distinction is further relevant since our histological analysis of synaptic puncta does not resolve between MLI:PC and PC:PC synapses, the relative balance of which may be altered by either KL or Kit manipulations. Indeed, future studies of the role of KL-Kit on PC-PC connectivity may inform how PC KL OX can apparently weaken inhibition to adjacent PCs: this may occur through the reduction of inhibitory drive from KL OX PCs to adjacent control PCs. An alternative, and not mutually exclusive possibility is that PC KL OX excessively potentiates MLI Kit to enhance MLI:MLI inhibition, reducing the inhibitory output surrounding MLIs can provide to control PCs.

Though such important nuances of the circuit mechanisms remain to be elucidated, our results demonstrate that the KL-Kit axis, from embryonic development through young adulthood, can alter GABAergic drive to PCs. These results suggest that KL-Kit may be necessary to maintain, and perhaps capable of modulating, MLI:PC GABAergic inhibition in the adult cerebellum. As each MLI (e.g. basket cell) has axon collaterals that can target multiple PCs, and as each PC can receive input from multiple MLIs, it is interesting to speculate that dynamics (e.g. in the expression or shedding) of KL may act through presynaptic Kit to tune the degree of convergent MLI:PC inhibitory drive. This would complement existing forms of modulation at the MLI:PC synapse, such as those dependent on other PC derived signaling molecules, like endocannabinoids [50-52] or peptides like secretin [53, 54] and CR [55]. The physiology of PCs, and the structure of MLI:PC inhibition, also varies over zones such as those identified by Zebrin-II and PLC β4 [56-60]; whether there is functional variation in KL or Kit expression, signaling, or dependency, remains to be determined. Though we have demonstrated a role in synaptic GABAergic inhibition, it is currently unknown whether Kit signaling also plays a role in other forms of inhibition in the cerebellar cortex, such as MLI:MLI gap-junctions, or MLI:PC ephaptic coupling. The latter may be suggested by the Kit KO mediated reduction in Kv1.1/Kv.12 and PSD-95 immunoreactivity.

While KL-Kit clearly influences synaptic function, we do not yet know the molecular mechanisms involved. Kit is a receptor tyrosine kinase that can activate multiple downstream effectors including Src family kinases, PLC, PI3K/mTOR, and the MAPK pathway, the latter two notable here for their shared affiliation with synapse phenotypes and autism spectrum disorder. Investigation of the interactome of Kit under basal and KL manipulated conditions may inform the receptor kinase cascades through which Kit signaling sustains synaptic function. It also plausible that trans-cellular interactions between KL and Kit may function in a kinase-independent, cell adhesion molecule modality [61, 62]. There is a substantial literature illustrating the importance of adhesion molecules in the organization and maintenance of MLI:PC inhibition, such as neurexins/neuroligins, dystrophin/dystroglycan, semaphorins, and neurofascins [44, 45, 63-67]. Much of what is understood about these molecules in the maintenance of PC inhibition is revealed by post-synaptic phenotypes; that is, cis-and trans interacting proteins that, via gephyrin, maintain the organization of PC GABAA receptors. It is notable therefore that MLI Kit or PC KL KO has a more predominant effect on mIPSC frequency than on amplitude, suggesting a presynaptic phenotype.

Despite the reduced PC inhibition resulting from MLI Kit KO or from PC KL KO, we have not observed overt cerebellar signs (i.e., ataxia). We suspect that this is due to compensation. Even the developmental ablation of functional GABAA channels in PCs does not produce overt motor signs, while adulthood ablation can [68]. Thus, we might unmask behavioral phenotypes if KL or Kit KO is delayed until adulthood. Consistent with the interpretation that developmental compensation may be involved, the impact of postnatal PC KL KO on inhibitory drive to PCs is more pronounced than embryonic MLI Kit KO. Beyond direct inhibition of PCs, MLI:PC GABAergic inhibition constrains granule cell mediated PC dendritic excitation and thus gates the conversion of PC LTD to LTP at the PF-PC synapse in motor learning [69, 70]. It is therefore possible that KL or Kit KO may only reveal phenotypes in motor learning (or other cerebellar-plasticity dependent) behaviors; speaking to the feasibility of this possibility, global hypomorphs for KL or Kit have altered hippocampal neurophysiology and learning performance on a Water Maze task [71-73]. Continuous expression of KL and Kit in specific neurons may thus reflect not only a role in the development (or maintenance) of synaptic function as suggested here, but also in the plastic modulation of synaptic strength.

Despite the importance of future works to determine molecular mechanisms and organismal consequences, our study is a discrete advance because it establishes that KL-Kit regulates mammalian central synapse function. Here, we demonstrate impacts at the mouse GABAergic MLI:PC synapse; however, Kit and KL exist in diverse organisms and neuronal populations including glutamatergic neurons of the hippocampus, and neurons in olfactory bulb, basal forebrain, and brainstem. It is therefore feasible that the synaptic phenotypes reported here reflect that KL-Kit is broadly capable of influencing connectivity.